---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Part I: Preprocessing and EDA
### Problem 1: Dataset Import & Cleaning

```{r}
## Part I: Preprocessing and EDA
#1.0
library(shiny)
library(dplyr)
library(tidyr)
library(ggplot2)

orders <- read.csv(file = "/Users/michaelwtarino/Desktop/Mike/Bootcamp/Week_7/Machine Learning Lab Part I & II/data/Orders.csv", stringsAsFactors =FALSE)
class(orders)
head(orders) 
dim(orders) #51290 rows by 24 columns
colnames(orders)

#Basic EDA
summary(orders)
sapply(orders, sd)
cor(orders)  #Error: 'x' must be numeric


### Problem 1: Dataset Import & Cleaning
class(orders$Profit)  ##factor and has $ dollar sign
class(orders$Sales)  ##factor and has $ dollar sign
orders$Profit <- gsub(",","",orders$Profit, fixed=TRUE)    #gets rid of commas
orders$Profit <- as.numeric(gsub("$","",orders$Profit, fixed=TRUE)) #gets rid of dollar sign and converts to numeric
orders$Sales <- gsub(",","",orders$Sales, fixed=TRUE)
orders$Sales <- as.numeric(gsub("$","",orders$Sales, fixed=TRUE))

#changes to numeric and gets rid of "," commas and $" dollar sign. 
orders$Order.Date
class(orders$Order.Date)
orders$Order.Date<-as.Date(orders$Order.Date, format = "%m/%d/%y") #lower case y
orders$Order.Date

```
### Problem 2: Inventory Management


```{r}
### Problem 2: Inventory Management
#Part 2.1
#plot by time
orders_q<-orders%>%
  group_by(.,Order.Date)%>%
  summarise(.,Quantity=sum(Quantity))

ggplot(data = orders_q , aes(x = Order.Date, y=Quantity))+
  geom_point(col="red",pch = 17)+
  xlab("Order.Date")+
  ylab("Quantity")+
  ggtitle("Relationship between Order.Date and Quantity")

#plot by month
orders_q_monthly <- orders %>%
  group_by(.,Order.Date= strftime(Order.Date,"%y/%m")) %>%
  summarize(.,Quantity = sum(Quantity))

ggplot(data = orders_q_monthly , aes(x = Order.Date, y=Quantity))+
  geom_point(col="red",pch = 17)+
  xlab("Order.Date")+
  ylab("Quantity")+
  ggtitle("Relationship between Order.Date and Quantity")


#Part 2.2
#plot by time and facet for category
orders_c<-orders%>%
  group_by(.,Order.Date, Category)%>%
  summarise(.,Quantity=sum(Quantity))

ggplot(data = orders_c , aes(x = Order.Date, y=Quantity))+
  geom_point(col="red",pch = 17)+
  xlab("Order.Date")+
  ylab("Quantity")+
  facet_wrap(~Category)+
  ggtitle("Relationship between Month and Quantity")


#plot by month and facet for category
orders_c_monthly <- orders %>%
  group_by(.,Order.Date= strftime(Order.Date,"%y/%m"), Category) %>%
  summarize(.,Quantity = sum(Quantity))

ggplot(data = orders_c_monthly , aes(x = Order.Date, y=Quantity))+
  geom_point(col="red",pch = 17)+
  xlab("Order.Date")+
  ylab("Quantity")+
  facet_wrap(~Category)+
  ggtitle("Relationship between Month and Quantity")

```
### Problem 3: Why did customers make returns?


```{r}
### Problem 3: Why did customers make returns?
returns <- read.csv(file = "/Users/michaelwtarino/Desktop/Mike/Bootcamp/Week_7/Machine Learning Lab Part I & II/data/Returns.csv", stringsAsFactors =FALSE)

#Merge the **Returns** dataframe with the **Orders** dataframe using `Order.ID`.
ordersandreturns <- merge(orders,returns,by="Order.ID")
ordersandreturns[is.na(ordersandreturns)] <- "No"
colnames(ordersandreturns)
head(ordersandreturns)

#3.1
# How much profit did we lose due to returns each year?
ordersandreturns_loss <- ordersandreturns %>%
  group_by(year = strftime(Order.Date, "%y")) %>% 
  filter(Returned == "Yes",Profit <0) %>%
  summarize(.,Loss = sum(Profit))
ordersandreturns_loss

#3.2
# How many customers returned 1,2,...etc times
ret_count <- ordersandreturns_custret %>%
  group_by(Returns) %>%
  summarize(Customers = n())
ret_count

#3.3
#Regions most likely to return orders
#Generate total returns by region
or_region_returns <- ordersandreturns %>%
  group_by(Region.x) %>%
  filter(Returned == "Yes") %>%
  summarize(Returns = n())

#Generate total orders by region
or_region_quantity = ordersandreturns %>%
  group_by(Region.x) %>%
  summarize(total = sum(Quantity))

#Merge both data frames 
r_d_q = merge(or_region_returns,or_region_quantity, by= "Region.x", all = TRUE)  
#Side note: merge(,by=,all = TRUE) keeps all rows and columns...
#...The default is FALSE, so that only rows with data from both x and y are included in the output.

#Divide total returns by total orders to get proportion or likelihood
r_d_q$proportion = r_d_q$Returns/r_d_q$total  #adds a proportion column to DF

#sort by greatest proportion of likelyhood
r_d_q[order(r_d_q$proportion, decreasing = TRUE),]

#3.4
#Categories / products most likely to be returned
#Generate total returns by sub-category
or_subs <- ordersandreturns %>%
  group_by(Sub.Category) %>%
  filter(Returned == "Yes") %>%
  summarize(Returns = n())

#Generate total orders by sub-category
or_subs_orders = ordersandreturns %>%
  group_by(Sub.Category) %>%
  summarize(total = sum(Quantity))

##Merge both data frames 
rdq = merge(or_subs_orders,or_subs, by= "Sub.Category", all = TRUE)

#Divide total returns by total orders to get proportion or likelihood
rdq$proportion = rdq$Returns/rdq$total    #adds a proportion column to DF

#sort by greatest proportion of likelyhood
rdq[order(rdq$proportion, decreasing = TRUE),]



```
## Part II: Machine Learning and Business Use Case
### Problem 4: Feature Engineering

```{r}
#4.1
#### Step 1: Create the dependent variable
#create a binary column named Returned1 to reflect returned orders where Yes = 1
ordersandreturns$Returned1 = as.numeric(ordersandreturns$Returned =="Yes")

#4.2
#### Step 2:
#crete a Process.Time column
class(orders$Ship.Date)
ordersandreturns$Ship.Date<-as.Date(ordersandreturns$Ship.Date, format = "%m/%d/%y") #lowercase y
ordersandreturns$Process.Time = ordersandreturns$Ship.Date - ordersandreturns$Order.Date

#4.3
#### Step 3:
#create a new data frame grouping by Product.ID - sum the 1s from the binary column to count the number of times returned.
returnedb4 = ordersandreturns %>%
  group_by(Product.ID) %>%
  summarize(returned.b4 = sum(Returned1))

#merge both dataframes to reveal the amount of orders returned per product.
ordersandreturns = merge(ordersandreturns,returnedb4, by= "Product.ID", all = TRUE)
ordersandreturns

```
### Problem 5: Fitting Models


```{r}
#remove unwanted columns and split data:
library(caret)

ordersandreturns_ml <- subset(ordersandreturns, select=-c(Product.ID,Order.ID,Row.ID,Order.Date, Customer.ID, Customer.Name)) #State too?
set.seed(123)
orIndex = createDataPartition(ordersandreturns$returned.b4, p=0.8, list = F, times = 1)
orTrain = ordersandreturns_ml[orIndex,]
orTest = ordersandreturns_ml[-orIndex,]

summary(ordersandreturns_ml)


# testing to make sure the ratios are correct
or_mltest <- ordersandreturns_ml %>%
  group_by(Ret_bin) %>%
  summarise(n())
or_mltest[2,2]/sum(or_mltest[,2])
#0.04328329
orTraintest <- orTrain %>%
  group_by(Ret_bin) %>%
  summarise(n())
orTraintest[2,2]/sum(orTraintest[,2])
#0.0438682
orTesttest <- orTest %>%
  group_by(Ret_bin) %>%
  summarise(n())
orTesttest[2,2]/sum(orTesttest[,2])
#0.04094365

# Fit a logistic regression model
install.packages("pROC")
library(pROC)

# Fitting a model
# Logistic regression
#model <- glm(Ret_bin ~.,
 #            family=binomial(link='logit'),
  #           data=orTrain)

```
### Problem 6: Evaluating Models


```{text}
##From Solutions##
If there is a product that has a high probability that will get returned, I would recommend my manager to put it on non-refundable sale so the customer could get it for a lower price but could not return the product. Alternatively, if the business is losing profit because we give a discount on a product, a model that has a relatively good accuracy with better precision (lower Type I error) is optimal.  

```
### Problem 7: Feature Engineering Revisit


```{text}



```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
